Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults

### 2024-01-30 manuscript_revised.RMD
finished revision

ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Knut Sindre Mølmen<https://orcid.org/0000-0001-8924-6848>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>

@@

**Purpose:** Resistance training-induced skeletal muscle hypertrophy seems to depend on ribosome biogenesis and -content. High glucose treatment may augment ribosome biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*. This was investigated with total RNA and ribosomal RNA as main outcomes, with relevant transcriptional or translational regulaters (c-Myc/UBF/rpS6) as a secondary outcome.

@@

Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardised training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013]. Genetic predisposition may explain some of this variation [@thalacker-mercer_cluster_2013], but in general the internal physiological milieu seems to be favourably shaped for muscle growth in response to different types of nutrient intake and exercise training [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013]. Indeed, exercise training and nutrients impact the ability to synthesise ribosomes, which in turn is demonstrated to be connected to the magnitude of RT-induced responses [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018]. For instance, increasing training volume generally induces greater ribosome biogenesis and is associated with greater benefits of RT in terms of gain in skeletal muscle mass and -strength [@hammarstrom_benefits_2020; @krieger_single_2009; @schoenfeld_strength_2017]. Still, as evident from Hammarström et al. (2020), not all participants experience increased muscle mass and improved muscle strength with increasing training volume. Therefore, means other than modification of RT variables seem necessary to optimise individual responses to RT, for instance nutritional adjuvants.

Nutritional supplements such as protein and creatine are frequently advocated as means to optimise RT adaptations [@cermak_protein_2012; @lanhers_creatine_2015; @lanhers_creatine_2017; @morton_systematic_2018]. The efficacy of other nutritional adjuvants such as glucose remains equivocal [@tezze_anabolic_2023]. This is surprising given that glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015; @sebata_control_2018]. Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn seems to determine muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019]. In addition, insulin, which is secreted from the beta cells of pancreas in response to rising blood glucose levels, may itself exert anabolic effects irrespective of muscle contraction, e.g. by elevating levels of amino acids, an also plays a role in reducing muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000]. It seems plausible that combined RT and glucose ingestion provides effects on ribosome biogenesis compared to RT alone.

@@

Sixteen healthy male and female participants (20-33 years, Table 1) were recruited to the study through social media advertisement and word of mouth. The eligibility criteria were non-smoking and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired muscle strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, while two participants experienced muscular discomfort related to heavy resistance training. Lean mass and body fat % (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

Table 1 could be placed here.

@@

All participants completed six RT sessions with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT. Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and plasma c-peptide levels). Participants were asked to avoid resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.

@@

Table 2 can be placed here.

@@

Glucose ingestion before and after RT sessions did not lead to a higher mean change of total training session volume, with a mean increase of 17% in both the glucose condition (pre: 5262 ± 1799kg, post: 6319 ± 2256kg, *p* > 0.05) and the placebo condition (pre: 5351 ± 1615kg, post: 6438 ± 2092, *p* > 0.05) from session 1 to session 6, respectively. There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* \> 0.05 for all, raw data and table available at GitHub repository).

@@

Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05). Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min, all : *p* \< 0.001). Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min: *p* = `r p.270`).

Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05). Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* \< 0.001).

@@

Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`). Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).

Both RT with glucose and placebo led to significantly reduced strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000). Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed an average increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`). Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`). Table 3 shows the mean change in absolute peak torque values per condition and angular velocity.

@@

The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% (glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1) and \~25-57% (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively), respectively (Figure 2). However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

@@

The five-session-RT intervention led to increases in all measured proteins both in the glucose and the placebo condition (Figure 3A). RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A). Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).

@@

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo. Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume. Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition. As by design, the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macronutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition. These findings suggest that the effects of high vs. low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design like the present study.

The observations of the present study from human skeletal muscle do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015]. Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000]. Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs. placebo conditions. Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT. Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis. While this is not a novel finding, it was only recently observed in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022]. As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

>>>>

ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Knut Sindre Mølmen <https://orcid.org/0000-0001-8924-6848>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>

@@

**Purpose:** Resistance training-induced skeletal muscle hypertrophy seems to depend on ribosome biogenesis and -content. High glucose treatment may augment ribosome biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*. This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome.

@@

Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardised training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013]. Genetic predisposition may explain some of this variation [@thalacker-mercer_cluster_2013], but in general, the internal physiological milieu seems to be favourably shaped for muscle growth in response to different types of nutrient intake and exercise training [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013]. Indeed, exercise training and nutrients impact the ability to synthesise ribosomes, which in turn is demonstrated to be connected to the magnitude of RT-induced responses [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018]. For instance, increasing training volume generally induces greater ribosome biogenesis and is associated with greater benefits of RT in terms of gain in skeletal muscle mass and -strength [@hammarstrom_benefits_2020; @krieger_single_2009; @schoenfeld_strength_2017]. Still, as evident from Hammarström et al. (2020), not all participants experience increased muscle mass and improved muscle strength with increasing training volume. Therefore, means other than modification of RT variables seem necessary to optimise individual responses to RT, for instance, nutritional adjuvants.

Nutritional supplements such as protein and creatine are frequently advocated as means to optimise RT adaptations [@cermak_protein_2012; @lanhers_creatine_2015; @lanhers_creatine_2017; @morton_systematic_2018]. The efficacy of other nutritional adjuvants such as glucose remains equivocal [@tezze_anabolic_2023]. This is surprising given that glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015; @sebata_control_2018]. Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn seems to determine muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019]. In addition, insulin, which is secreted from the beta cells of the pancreas in response to rising blood glucose levels, may itself exert anabolic effects irrespective of muscle contraction, e.g. by elevating levels of amino acids, and also plays a role in reducing muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000]. It seems plausible that combined RT and glucose ingestion provide additive effects on ribosome biogenesis compared to RT alone.

@@

Sixteen healthy male and female participants (20-33 years, Table 1) were recruited to the study through social media advertisement and word of mouth. The eligibility criteria were non-smoking and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired muscle strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, while two participants experienced muscular discomfort related to heavy resistance training. Lean mass and body fat % (Table 1) were measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

*Table 1 could be placed around here.*

@@

All participants completed six RT sessions with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT. Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and serum c-peptide levels). Participants were asked to avoid resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimise interference from external exercise sources.

@@

*Table 2 can be placed here.*

@@

Glucose ingestion before and after RT sessions did not lead to a higher mean change of total training session volume, with a mean increase of 17% in both the glucose condition (pre, 5262 ± 1799kg; post, 6319 ± 2256kg, *p* > 0.05) and the placebo condition (pre, 5351 ± 1615kg; post, 6438 ± 2092, *p* > 0.05) from session 1 to session 6, respectively. There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* \> 0.05 for all, raw data and table available in GitHub repository).

@@

Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0min, 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min, 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min, 1.62 ± 1.10 mmol/L, all *p* < 0.001), with no changes being observed in the placebo condition (Figure 1B, 0 min, 0.09 ± 0.3 mmol/L; 15 min, 0.16 ± 0.35 mmol/L; 30 min, 0.18 ± 0.39 mmol/L, all *p* > 0.05). Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min, all *p* \< 0.001). Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min, *p* = `r p.270`).

Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min, 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min, 793 ± 581.0 pmol/L, both *p* < 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min, 63.7 ± 71.0 pmol/L; 30 min, 53.9 ± 134.0 pmol/L; both *p* > 0.05). Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* \< 0.001).

*Figure 1 can be placed here.*

@@

In general, glucose ingestion before and after RT sessions did not improve skeletal muscle recovery compared to placebo throughout the intervention, neither the rested state (Figure 1D, 23hrs after exercise, Post 2RT, *p* = `r pdiff.c2`; Post 4RT, *p* = `r pdiff.c3`), nor acutely after the sixth/final RT-session (30min post 6RT, *p* = `r pdiff.c5`; 2h post 6RT, *p* = `r pdiff.c6`) or in the rested state after the sixth RT session (23h post 6RT, *p* = `r pdiff.c7`). In contrast to this, glucose ingestion was associated with a 7% less reduction in muscle strength after the fifth session compared to placebo (Figure 1D, *p* = `r pdiff.c4`).

In knee extension torque, both RT with glucose and placebo led to significantly reduced muscle strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT, *p* = 0.000). Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed an average increase in muscle strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT, *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT, *p* = `r apc3`). Twenty-three hours after the last (sixth) RT session, muscle strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT, *p* = `r apc4`). Table 3 shows the mean change in absolute peak torque values per condition and angular velocity.

*Table 3 can be placed here.*

@@

The five-session-RT intervention led to on average ~20-27% increases in total RNA (glucose, 263 ± 50 ng/mg^-1^; placebo, 210 ± 121 ng/mg^-1^) and ~25-57% increases in rRNA per unit muscle weight (47S, 0.253 ± 1.27 and 0.576 ± 0.677; 18S, 0.336 ± 0.460 and 0.271 ± 0.470; 28S, 0.314 ± 0.504 and 0.311 ± 0.582; 5.8S, 0.388 ± 0.576 and 0.322 ± 0.520; 5S, 0.305 ± 0.608 and 0.292 ± 0.432; arbitrary units for glucose and placebo respectively) (Figure 2). However, RT with glucose did not induce more pronounced accumulation of total RNA compared to RT with placebo (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`).

*Figure 2 can be placed here. *

@@

The five-session-RT intervention led to increased abundances of all measured proteins, both in the glucose and in the placebo condition (Figure 3A). RT with glucose resulted in lowered estimates of c-Myc, UBF and rpS6 levels compared to placebo (being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively), without reaching statistical significance (Figure 3A, *p* = `r min(p_west)`-`r max(p_west)`). Baseline and trained-state total RNA levels showed a linear relationship with UBF abundances; a $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 SD unit increase in UBF (Figure 3C, *p* = `r corrp`).

*Figure 3 can be placed here.*

@@

The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis compared to RT with placebo, measured as total RNA, rRNA, and protein involved in rDNA transcription initiation. Similarly, RT with glucose  did not affect markers of skeletal muscle functionality compared to placebo, such as muscle strength and recovery, and total training session volume. Towards the end of the intervention, RT with glucose led to less reduction in muscle strength compared to placebo, which may point to an accumulated effect of the glucose condition. As by design, the levels of plasma glucose and serum c-peptide were significantly higher before, during and after the glucose condition RT sessions compared to placebo, and there were no differences in daily macronutrient intake between conditions on consecutive days, suggesting that the study was adequately designed to elucidate the biological and functional effects of the glucose condition. These findings suggest that the previously observed significance of glucose exposure for rDNA transcription initiation in *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] is not translatable to acute effects in resistance exercised human skeletal muscle *in vivo*, with a design like the present study.

The observations of the present study from human skeletal muscle cells do not support previous indications of high glucose exposure on rDNA transcription initiation in non-human and/or non-skeletal muscle cells [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. A UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015]. Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000]. Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022]. Therefore, in the present study, it was expected that any benefits of ingesting glucose compared to placebo with RT would be measurable after five training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012]. Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study displayed no signs of such effects of glucose vs. placebo conditions.

Importantly, previous studies investigated high vs. low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs. glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition (stevia), with a matched daily macronutrient and energy intake. Therefore, the comparison made in the present study was high plasma glucose levels vs. normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status. Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting added glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT. Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells. Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].

In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc were performed. These analyses further supported observations from the RNA data as there was no difference between the glucose and placebo conditions. UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 previously correlated with 18S and 28S rRNA content and is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016]. Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016]. Hence, it seems quite reasonable to observe similar changes in these three proteins. The linear relationship found exclusively between UBF content and total RNA levels, and not between total RNA and c-Myc or rpS6, supports a specific role for UBF in regulating ribosomal content in human skeletal muscle. This is in line with recent observations in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022]. The observed RT-induced increase in levels of UBF, c-Myc and rpS6 is itself in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012].

As in the biological data, combined glucose ingestion and RT did not exert measurable effects on muscular strength throughout the intervention compared to placebo. In general, the skeletal muscle performance index, which was used as a proxy marker for muscular recovery, decreased similarly from baseline to after the intervention in both conditions. There was one exception to this however, as glucose ingestion was associated with a lower reduction in muscular strength after five RT sessions compared to placebo, which may point towards a beneficial accumulated effect where the heavy-load RT gradually fatigued the participants but glucose ingestion counteracted this response [@mul_exercise_2015; @sebata_control_2018]. Therefore, we cannot rule out potential long-term benefits of ingesting glucose in connection with heavy-load resistance training, which would require a longer intervention period than that of the present study. Having noted this, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measured 30 minutes, 2 hours and 23 hours after the sixth/final training session. As such, the potential accumulated effect observed after five RT sessions did not extend to acute effects measured after RT session six.

The decrease in muscular performance observed over the course of the intervention might be explained by the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise. Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000]. Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30 minutes and 2 hours after the sixth RT session and a drop at 23 hours after the sixth session. Further, muscle strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests. However, this neither explains the difference observed after five RT sessions between conditions, nor the drop in muscle strength from after the fourth RT session to after the fifth RT session. A possible explanation could be that exercising without glucose ingestion may have caused more lower cellular energy substrate availability compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo. Notably, training volume data displayed that the total training session volume was not different on pairwise consecutive days, i.e. no difference between days 1-2, days 3-4 and so on. Hence, there were no differences in mechanical loading to induce greater fatigue between conditions. Arguably, an increased energy availability via glucose ingestion during RT may induce less acute fatigue on the exercised skeletal muscle, and perhaps less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999]. Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999]. Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
